---
title: "Data Visualization"
author: "Michael Zargari"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages

```{r}
# General use packages:
library(tidyverse)
library(here)
library(janitor)

# Specifically for plots:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

# And for another dataset we'll explore:
library(gapminder)

# Spatial
library(sf)
```

### Read in Lizard data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))

lizard_counts <- lizards %>% 
  group_by(common_name) %>% 
  summarize(
    counts = n()
  )

## This does the same thing:
lizard_counts2 <- lizards %>% 
  count(common_name, site, tail)

lizards_median <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))
```

### Plot them

```{r}
ggplot(data = lizards, aes(x = total_length     ## Makes changes global changes
                           , y = weight)) +
  geom_point()                                  ## Putting any code in here only affects the line
```

Create a histogram of total lengths for all lizards in the dataset

```{r}
## aes should only contain variables, not constants

ggplot(data = lizards, aes(x = total_length)) +
  geom_histogram(color = "red"          # changes the border color
                 , fill = "blue"        # changes the fill color
                 , size = 1             # changes the size of the outline
                 , linetype = "dotted") # changes the type of outline
```

```{r}
ggplot(data = lizards, aes(x = total_length   
                           , y = weight)) +
  geom_point(shape = 22             ## Each shape number is a different shape
             , color = "yellow"
             , fill = "red"
             , size = 4
             , alpha = .2)          ## This controls how see-through the points are  
```

```{r}
ggplot(data = lizards, aes(x = total_length   
                           , y = weight)) +
  geom_point(aes(color = common_name
                 , size = total_length)
             , alpha = 0.5) +
  theme_minimal()
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  facet_wrap(~common_name, ncol = 3, scales = "free")
```

```{r}
ggplot(data = lizard_counts, aes(x = counts
                                 , y = fct_reorder(common_name, counts))) +
  geom_col()
```

Create a scatterplot, faceted by species for total length versus weight. The color and shape of points should differ based on the species (common name). Change the theme. Make each faceted panel have flexible axis limits (scales should be free)

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = weight)) +
  geom_point(aes(color = common_name)) +
  theme_light() +
  facet_wrap(~common_name, scales = "free") +
  labs(x = "Total Length (mm)"
       , y = "Weight (grams)"
       , title = "Jonada Basin lizard sites"
       , subtitle = "This is a subtitle"
       , caption = "Data is from The Organization")
```

### Fun with scales

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = weight)) +
  geom_point(aes(color = weight)) +
  scale_colour_gradient(low = "green", high = "black")
```

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = weight)) +
  geom_point(aes(color = weight)) +
  scale_colour_gradientn(colors = c("magenta", "cyan4", "darkorchid3", "dodgerblue"))
```

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = weight)) +
  geom_point(aes(color = weight)) +
  scale_colour_steps(low = "yellow", high = "black")
```

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_colour_steps2(low = "yellow"
                      , mid = "orange"
                      , high = "black"
                      , midpoint = 150
                      , breaks = c(50, 75, 150, 180, 220, 280)) + 
  theme_minimal()
```

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("magenta", "cyan4", "darkorchid3", "dodgerblue")
                     , breaks = seq(from = 0, to = 300, by = 60))
```

Make a horizontal boxplot with commons name on the y-axis, total length on the x-axis, with boxplot fill changing based on the common name

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = common_name)) +
  geom_boxplot(aes(fill = common_name))
```

This orders it by the median of the group

```{r}
ggplot(data = lizards_median, aes(x = total_length
                         , y = common_name)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "rcartocolor::Fall") +
  theme_minimal()
```

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = weight)) +
  geom_point(aes(color = weight)) +
  theme(panel.grid.major.x = element_line(color = "red")
        , panel.grid.minor.y = element_line(color = "green")
        , axis.title.x = element_text(color = "blue", size = 12)
        , axis.title.y = element_text(color = "orange")
        , text = element_text(size = 18)
        , panel.background = element_rect(color = "purple", fill = "pink")) +
  annotate("text", x = 300, y = 50, label = "This is the point\nwe made", color = "white") +
  geom_vline(xintercept = 150, linetype = "dashed", color = "orange", size = 3)
```

## Repulsive labels with `ggrepel`

```{r}
wws_lizards <- lizards %>% 
  filter(common_name == "western whiptail"
         , site == "sand")
```

```{r}
ggplot(data = wws_lizards, aes(x = total_length
                         , y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num))
```

```{r}
gapminder %>% 
  filter(year == 2002 & continent == "Europe") %>% 
  ggplot(aes(x = gdpPercap
             , y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)
```

```{r}
ggplot(data = lizards, aes(x = total_length
                         , y = weight)) +
  geom_point() +
  gghighlight(toe_num == 250, label_key = toe_num)
```

### Making a map

```{r}
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot() + 
  geom_sf(data = jornada_veg, aes(fill = Name)) + 
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
```
